### 2023-10-24 exploreation.Rmd
first commit

```{r}

N <- 100
ki  <-  3
km  <-  1
K  <- ki+km
ntraits=list(a=c(
etopo <- read.csv(textConnection(
"altitudes,colours
10000,#FBFBFB
4000,#864747
3900,#7E4B11
2000,#9B8411
1900,#BD8D15
300,#F0CF9F
0,#307424
-1,#AFDCF4
-12000,#090B6A
"
), stringsAsFactors=FALSE)
etopo$altitudes01 <- scales::rescale(etopo$altitudes)

etopo.colors <- function(n) {
  colorRampPalette(etopo$colours)(n)
  }

z=4
a=rep(0,3)
adaptiveTraits=a
age  <- vector()
community  <- vector()
population  <- cbind(age=age,community=community)

community <- list(
                  coordinates=cbind(x=vector(mode="numeric",length=K),y=vector(mode="numeric",length=K)),
                  traits=cbind(a1=vector(mode="numeric",length=K),a2=vector(mode="numeric",length=K),a3=vector(mode="numeric",length=K)),
                  size=vector(mode="numeric",length=K)

                  )
source("R/population.R")
N=100
z=4 #number of neutral traits
ki <- 5
km <- 2
K  <-  ki+km
m=.2 #proba marriage
b=.1 #HG birth rate
r=.2 #farmer scale param
rho #marriage rule

a=initAdaptiveTraits(ki=ki,km=km)
neutraltraitsParam=initNeutralTraits(z = z)
neutraltraitsParam$s=c(0,0,1,1)

reproduction <- function(p1,p2,tp,i=NULL){
    print(i)
    stopifnot(p1$community==p2$community)
    traits=ifelse(sample(c(0,1),length(tp$s),replace=T),p1$traits,p2$traits)
    list(
         id=i,
         traits=traits,
         age=0,
         community=p1$community,
         sex=sample(c("F","M"),1),
         repro=FALSE,
         partner=-1
    )
}

population=lapply(1:N,function(i)
                      list(
                           id=i,
                           traits=sample(c(0,1),size=z,replace=T),
                           age=sample.int(80,1),
                           community=sample.int(K,1),
                           sex=sample(c("F","M"),1),
                           repro=FALSE,
                           partner=-1
                      )
)

pos=random2Dgrid(K=K,Gx=100)
initcomus=initialiseCommunities(traits=a,coordinates=pos)
size=table(sapply(population,"[[","community"))
initcomus$size=size
plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))

tstep=1000
popsize=c()
popsum=list()
for(time in 1:tstep){
    for(i in sample(1:length(population))){
        if(!is.null(population[[i]])){
            ind=population[[i]]$id
            ##ageing
            population[[ind]]$age=population[[ind]]$age+1

            ##repro
            partner=population[[ind]]$partner
            repro=population[[ind]]$repro
            if(partner>0 & !repro){
                print(paste("childtest",ind,partner,population[[ind]]$community,population[[partner]]$community))
                population[[partner]]$repro=TRUE #make sure proba of reproduction is unique for all pair 
                ad_tr=initcomus$adaptivetraits[population[[ind]]$community,]
                lambda=b+r*sum(ad_tr)
                nchild=rpois(1,lambda)
                if(nchild>0){
                    for(i in 1:nchild){
                        print("newborn")
                        population[[length(population)+1]]=reproduction(population[[ind]],population[[partner]],neutraltraitsParam,length(population)+1)
                    }
                }
            }
            population[[ind]]$repro=FALSE

            ##marriage
            if(population[[ind]]$age>18 & population[[ind]]$partner <0){
                print("hmmmaybe")
                if(runif(1)<m){
                    potential=sapply(population,function(i,s)if(i$age>18 & i$partner<0 & i$sex != s )return(i$id),s=population[[ind]]$sex)
                    potential=unlist(potential)
                    if(length(potential)>0){
                        if(length(potential)==1)partner=potential
                        else partner=sample(potential,1)
                        population[[ind]]$partner=partner
                        population[[partner]]$partner=ind
                        newcom=ifelse(runif(1)<rho,population[[ind]]$community,population[[partner]]$community)
                        population[[ind]]$community=population[[partner]]$community=newcom
                        print(paste("marriage",ind,partner,"moving all to",newcom,":",population[[partner]]$community,population[[ind]]$community))
                    }
                }
            }
            #if(runif(1)<d)population[[ind]]=NULL
        }

        ##pm transmission

        ##pm transmission
        #quickly summarize
        popsum[[time]]=sapply(names(population[[1]]),function(n)table(sapply(population,"[[",n)))
        size=table(sapply(population,"[[","community"))
        initcomus$size=size
        ##
        plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))
        popsize=c(popsize,length(population))
    }
}

plot(popsize)

sumcomu=sapply(popsum,"[[","community")
plot(1,1,ylim=range(0,sumcomu),xlim=c(0,ncol(sumcomu)),type="n")
lapply(1:K,function(k) lines(sumcomu[k,],col=mean(initcomus$adaptivetraits[k,])+1,lwd=2) )