### 2024-01-12 inst/scripts/fissionExplore.Rmd

```{r}
devtools::load_all(".")
N=400
Th=400 #fission threshold
ki <- 1
km <- 1
K  <-  ki+km
m=1 #proba marriage
b=.2 #HG birth rate
r=.05 #farmer scale param
rho=.5 #marriage rule
d=0.001
maturity=0
endrepro=20
tstep=20
z=5

neutraltraitsParam=generatePathways(z = z)
neutraltraitsParam$s=c(0,0,1,1)
neutraltraitsParam$pre[,"v"]=c(1,1,1,0,0)

pos=random2Dgrid(K=K,Gx=100)
a=initAdaptiveTraits(ki=ki,km=km)
initcomus=initialiseCommunities(traits=a,coordinates=pos)
initcomus$size=rep(round(N/K),K)
if(sum(initcomus$size) != N){
    ndiff=N-sum(initcomus$size)
    if(ndiff>0){
        ss=sample(1:K,ndiff,replace=TRUE)
        for(i in ss) initcomus$size[i]=initcomus$size[i]+1
    }
    if(ndiff<0){
        ss=sample(1:K,-ndiff,replace=TRUE)
        for(i in ss) initcomus$size[i]=initcomus$size[i]-1
    }
}

plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))

communities=unlist(lapply(1:K,function(i)rep(i,initcomus$size[i])))
population=cbind(newpop(N,age="random",community = communities),generateTraitsMatrix(N,z))

quickV=modelVector( K =K, m=m, b=b, r=r, rho=rho, d=d, maturity=maturity, endrepro=60, population=population, comus=initcomus, tstep=10, tp=neutraltraitsParam, logging=c("time"),ma=1,traitsid=paste0("t",1:z),getfinalpop=TRUE,F_Th=10800,out=c("finalpop","popsize","weddings"))
hist(quickV$population[,"age"])
hist(table(quickV$population[,"fid"]))
testpop=quickV$population
tapply(testpop[,c("fid")],testpop[,"community"],table)

# using forloop
fmig=c()
fids=sample(unique(testpop[testpop[,"community"] ==3,"fid"]))
lim=1
fmig=fids[1:lim]
while(sum(testpop[,"fid"]%in% fmig)<800){
          lim=lim+1
          fmig=fids[1:lim]
}

# using cumsum
compop <- testpop[testpop[, "community"] == 3, ]
fids <- compop[, "fid"]
cumulative_size <- cumsum(table(fids)[as.character(sample(unique(fids)))])
selected_families <- names(cumulative_size[cumulative_size < 800])

##comparison
microbenchmark::microbenchmark(reassignFamiliesToNewCommunity(2,testpop,F_Th,length(initcomus$size)+1),splitCommunitiesByFamilies(2,testpop,F_Th,length(initcomus$size)+1))

a=splitCommunitiesByFamilies(2,testpop,F_Th,length(initcomus$size)+1)
table(a[,"community"])
a=reassignFamiliesToNewCommunity(2,testpop,F_Th,length(initcomus$size)+1)
unique(a[a[,"community"]==length(initcomus$size)+1,"fid"])

alltFL=replicate(100,table(splitCommunitiesByFamilies(2,testpop,F_Th,length(initcomus$size)+1)[,"community"]))
alltCS=replicate(100,table(reassignFamiliesToNewCommunity(2,testpop,F_Th,length(initcomus$size)+1)[,"community"]))
alltFL[2,] >= alltFL[4,]
alltCS[2,] >= alltCS[4,]

##cumsum is 6/7 time faster

quickV=modelVector( K =K, m=.5, b=0.03*3.1, r=0, rho=rho, d=0.03, maturity=maturity, endrepro=65, population=population, comus=initcomus, tstep=tstep, tp=neutraltraitsParam, logging=c("time","visu","fission","marriage"),ma=1,traitsid=paste0("t",1:z),out=c("popsumary","finalpop","popsize","weddings"),F_Th=800)
tstep=1000

quickV=modelVector(m=1, b=0.03*6, r=0, rho=rho, d=0.03, maturity=18, endrepro=65, population=population, comus=initcomus, tstep=200, tp=neutraltraitsParam, logging=c("time","visu","fission","marriage"),ma=1,traitsid=paste0("t",1:z),out=c("popsumary","finalpop","popsize","weddings"),F_Th=800)

par(mfrow=c(1,2))
hist(quickV$population[,"age"],xlab="age")
hist(table(quickV$population[,"fid"]),xlab="family size")
testpop=quickV$population
tapply(testpop[,c("fid")],testpop[,"community"],table)

quickV=modelVector( K =K, m=m, b=b, r=r, rho=rho, d=d, maturity=maturity, endrepro=60, population=population, comus=initcomus, tstep=10, tp=neutraltraitsParam, logging=c("time"),ma=1,traitsid=paste0("t",1:z),getfinalpop=TRUE,F_Th=800,out=c("finalpop","popsize","weddings"))

population=quickV$population

table(population[,"community"])

reassignFamiliesToNewCommunityNoFIDs(2,population,200,3)

quickV=modelVector( K =K, m=m, b=b, r=r, rho=rho, d=d, maturity=maturity, endrepro=60, population=population, comus=initcomus, tstep=100, tp=neutraltraitsParam, logging=c("time"),ma=1,traitsid=paste0("t",1:z),getfinalpop=TRUE,F_Th=10800,out=c("finalpop","popsize","weddings"))
```

Using new fissionning:

```{r}
devtools::load_all(".")
N=400
ki <- 1
km <- 1
K  <-  ki+km
z=5

neutraltraitsParam=generatePathways(z = z)
neutraltraitsParam$s=c(0,0,1,1)
neutraltraitsParam$pre[,"v"]=c(1,1,1,0,0)

pos=random2Dgrid(K=K,Gx=100)
a=initAdaptiveTraits(ki=ki,km=km)
initcomus=initialiseCommunities(traits=a,coordinates=pos)
initcomus$size=rep(round(N/K),K)

communities=unlist(lapply(1:K,function(i)rep(i,initcomus$size[i])))
population=cbind(newpop(N,age="random",community = communities),generateTraitsMatrix(N,z))
quickV=modelVector(m=1, b=0.03*6, r=0.012, rho=.5, d=0.03, maturity=18, endrepro=65, population=population, comus=initcomus, tstep=500, tp=neutraltraitsParam, logging=c("time","visu","fissionsize"),ma=1,traitsid=paste0("t",1:z),out=c("popsumary","finalpop","popsize","weddings"),F_Th=200)
```